---
title: "Analysis of HH Without Children"
author: "Martha Moreno"
date: "October 4 2017"
output: github_document
---

```{r setup, include=FALSE}

rm(list=ls())

knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(ltm)
library(foreign)
library(ggplot2)
library(irtoys)
library(gmodels)
library(mirt)

#require(tidyverse)

#setwd ("C:/Users/Martha/Documents/Applied Statistics/3rd semester/Independent Study")
opts_knit$set(root.dir = "C:/Users/Martha/Documents/Applied Statistics/3rd semester/Independent Study")

```


#Research Question:
Fitting several IRT  models on the Food Security Scale to assess:

1) Which model has better fit between: 
    a) unconstrained Rasch Model
    b) constrained Rasch Model
    c) Two Parameter Logistic Model

2) Can we obtain significantly different cut-off points using the best model?

2) If the model shows a good fit, do we have parameter invariance between Mexico압 and U.S. scales?

##Food Security Scale
Data from the Current Population Survey - December of 2016

##How is the survey conducted?

####Screening question - not used for food security measure and so not used for this analysis:

 - HESS1 = i1 

Which of these statements best describes the food eaten in your household in the last 12 months: we always have enough to eat and the kinds of food we want; we have enough to eat but not always the kinds of food we want; sometimes we don't have enough to eat; or often we don't have enough to eat?

--------------------------------------------------------

####Stage 1

 - HESS2 =	i2

Scale12 - worried food would run out, past 12 months	

With the following Ranges:	

-9  No response	

-3  Refused	

-2  Don't Know	

1  Often true	

2  Sometimes true	

3  Never true  

-------------------------------------------------------

 - HESS3 = i3

Scale12 - food bought didn't last, past 12 months

With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Often true

2  Sometimes true

3  Never true  

-------------------------------------------------------
 
 - HESS4 = i4

Scale12 - couldn't afford balanced meals, past 12 months

With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Often true

2  Sometimes true

3  Never true  

------------------------------------------------------- 

####If there are children <18 ask q5 and q6  

 - HESS5 = i5

Scale12 - child fed few kinds of low-cost foods, past 12 months
With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Often true

2  Sometimes true

3  Never true  

------------------------------------------------------- 

 - HESS6 = i6

Scale12 - couldn't feed child balanced meals, past 12 months
With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Often true

2  Sometimes true

3  Never true  

-------------------------------------------------------

####1st-level Screen (screener for Stage 2): If AFFIRMATIVE RESPONSE to ANY ONE of Questions 2-6 (i.e., "often true" or "sometimes true") OR response [3] or [4] to Question 1 (if administered), then continue to Stage 2; otherwise, skip to end.

####stage 2

 - HESH1 = i7

Scale12 - child not eating enough, past 12 months

With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Often true

2  Sometimes true

3  Never true  

------------------------------------------------------- 

 - HESH2 = i8

Scale12 - adult cut size or skipped meals, past 12 montths
With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Yes

2  No  

------------------------------------------------------- 

 - HESHF2 = i8a

Scale12 - frequency adult cut or skipped, past 12 mon
With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Almost every month

2  Some months but not every month

3  Only 1 or 2 months

------------------------------------------------------- 

 - HESH3 = i9

Scale12 - ate less than felt they should, past 12 months

With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Yes

2  No

------------------------------------------------------- 
 
 - HESH4 = i10

Scale 12 - respondent hungry but didn't eat, past 12 months

With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Yes

2  No

------------------------------------------------------- 

 - HESH5 = i11

Scale12 - respondent lost weight, past 12 months With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Yes

2  No

-------------------------------------------------------

#### 2nd-level Screen (screener for Stage 3): If AFFIRMATIVE RESPONSE to ANY ONE of Questions 7 through 11, then continue to Stage 3; otherwise, skip to end.

#### Stage 3

 - HESSH1 = i12

Scale12 - adult not eat whole day, past 12 months With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Yes

2  No

-------------------------------------------------------

 - HESSHF1 = i12a

Scale12 - frequency adult not eat whole day, past 12 months
With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Almost every month

2  Some months but not every month

3  Only 1 or 2 months

-------------------------------------------------------

 - HESSH2 = i13
Scale12 - cut size of child's meals, past 12 months With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Yes

2  No

-------------------------------------------------------

 - HESSH4 = i14

Scale12 - child skipped meal, past 12 months With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Yes

2  No
-------------------------------------------------------

 - HESSHF4 = i14a

Scale12 - frequency child skipped meal, past 12 months With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Almost every month

2  Some months but not every month

3  Only 1 or 2 months

------------------------------------------------------- 

 - HESSH3 = i15

Scale12 - child hungry - past 12 months With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Yes

2  No

------------------------------------------------------- 

 - HESSH5 = i16

Scale12 - child didn't eat whole day, past 12 months With the following Ranges:

-9  No response

-3  Refused

-2  Don't Know

1  Yes

2  No

-------------------------------------------------------

####Some considerations about the data:

- For the often true/sometimes true/never true items, they are recorded as follows:

recoded as 1 if "Often true" or "Sometimes true" and 0 if "Never true" 

- For the yes/no items, they are recorded as follows: 

recoded as 1 if "Yes" and 0 if "No" 

- For the almost every month/some months but not every month/only 1 or 2 months items they were recoded as follows: 

recoded as 1 if "every month" or "some months but not every month" and 0 if "only 1 or 2 months" 


- The rawscore that comes along the data set from the survey does not match my generated rawscore in 264 cases, I decided to exclude these cases from the analyisis until I figure out what압 going on here.

- The households with no information are also excluded from the analysis (29,192 cases ehich represent 22.6% of the whole sample)


```{r}

dat<-read.dta("data/data_clean_nochild.dta")

items1=cbind(dat$i2, dat$i3, dat$i4, dat$i8, dat$i8a, dat$i9, dat$i10, dat$i11, dat$i12, dat$i12a)
items2=cbind(dat$i2, dat$i3, dat$i4, dat$i8, dat$i9, dat$i10, dat$i11, dat$i12)

#Frequencies
freq1<-apply(items1,2,table)
freq2<-apply(items2,2,table)
freq1<-addmargins(freq1)
freq2<-addmargins(freq2)

#Means of affirmative answers
means1<-apply(items1,2,mean)
means2<-apply(items2,2,mean)

#Summary table of 10 items
table1<-rbind(round(freq1[1:3,1:10],0),round(means1*100,2))
nam1<-c("Item2", "Item3", "Item4", "Item8", "Item8a", "Item9", "Item10", "Item11", "Item12", "Item12a")
nam2<-c("No", "Yes", "Total", "Percentage")
colnames(table1)<-nam1
rownames(table1)<-nam2

#Summary table of 8 items
table2<-rbind(round(freq2[1:3,1:8],0),round(means2*100,2))
nam1<-c("Item2", "Item3", "Item4", "Item8", "Item9", "Item10", "Item11", "Item12")
nam2<-c("No", "Yes", "Total", "Percentage")
colnames(table2)<-nam1
rownames(table2)<-nam2
colnames(items2)<-nam1

#Plots of affirmative answers for each item
means.plot1<-qplot(x=1:length(means1), y=means1, main="Proportion of Affirmative Answers 10 items", xlab="Items for hh Without Children", ylab="") + geom_line()
means.plot2<-qplot(x=1:length(means2), y=means2, main="Proportion of Affirmative Answers 8 items", xlab="Items for hh Without Children", ylab="") + geom_line()

#Using descript from ltm to get the biserial correlations:
d1<-descript(items1)
correlations1 <- d1$bisCorr
d2<-descript(items2)
correlations2 <- d2$bisCorr

#Biserial correlations plots
cor.plot1<-qplot(x=1:length(correlations1), y=correlations1, main="Biserial Correlations 10 items", xlab="Items for hh without children", ylab="") + geom_line()
cor.plot2<-qplot(x=1:length(correlations2), y=correlations2, main="Biserial Correlations 8 items", xlab="Items for hh without children", ylab="") + geom_line()

d1
d2


```

####Descriptive analysis for households without children (8 items)

####Frequencies and means

```{r}

table1
table2

means.plot2
means.plot1

```

####Biserial correlations

```{r}

correlations1
correlations2

cor.plot1
cor.plot2

```

####Status of food security vs raw score

```{r}

CrossTable(dat$rawscore,dat$status, expected = F, prop.r=F, prop.c=F, prop.t=F, prop.chisq=F, chisq = F, fisher=F, mcnemar=F, resid=F, sresid=F, asresid=F)

```

####Testing several models just using the 8 items (no follow ups)

1) Fit a Rasch Model. Rasch models are also known as one parameter models. It assumes that the guessing and the discrimination parameter are the same across items:

In this case we are fitting an unconstrained and a constrained Rasch Model with:

  $\alpha_i=2$ $\forall i$ ??? - this was just for the graphs, but how do you establish it??
  
  $c_i=0$ $\forall i$

2) Fit a Two Parameter Model, in this case the model estimates the discrimination parameter for each item, but $c_i=0$ $\forall i$

####Rasch Model

```{r}

#Fit a Rasch model - hessian matrix is non-positive!!! None of these work
#fit.rasch<-rasch(items2)
#summary(fit.rasch)
#fit.rasch <- tpm(items2, type = "rasch", constraint = cbind(1:8, 3, 1))
#summary(fit.rasch)

#Constraint? - this fits better? In order to impose the constraint alpha, the constraint argument is used. This argument accepts a two-column matrix where the first column denotes the parameter and the second column indicates the value at which the corresponding parameter should be fixed.
fit.rasch<-rasch(items2, constraint=cbind(8+1,1))
summary(fit.rasch)

#Margins?? - we don앖 want앖 the chi-square to be significant
margins(fit.rasch)
margins(fit.rasch, type="three-way")

#Goodness of fit cannot be computed with unconstrained model?? - null hypothesis states that the observed data have been generated under the Rasch model with the parameter values the maximum likelihood estimates.
#GoF.rasch(fit.rasch, B = 100)

#Residuals
#plot residuals vs ability to test homoscedasticity??
res.rasch<-residuals(fit.rasch)

#Factor scores
fs.rasch<-factor.scores(fit.rasch)
#plot(fs.rasch)

#ICC and information curves
par(lwd=2)
plot(fit.rasch, legend = TRUE, cx = "left", cex.main = 1.5, cex.lab = 1.5, cex = 1.5)
plot(fit.rasch, type = "IIC", annot = FALSE, cex.main = 1.5, cex.lab = 1.5)
plot(fit.rasch, type = "IIC", items = 0, cex.main = 1.5, cex.lab = 1.5)

```




####2PL Model

```{r}

#Fit a 2PL model - up to two latent variables with names z1 and z2
#I changed the number of quadrature points in order to get coefficients more similar to the mirt package??
fit.2pl<-ltm(items2~z1, IRT.param = T, control=list(GHk = 61, iter.qN = 50))

summary(fit.2pl)

#Margins - is this due to sample size??
margins(fit.2pl)
margins(fit.2pl, type="three-way")

#Compare the fit of the Rasch Model vs the 2PL Model - is this a better fit??
anova(fit.rasch, fit.2pl)

#Residuals
#plot residuals vs ability to test homoscedasticity??
res.2pl<-residuals(fit.2pl)

#Factor scores - is the lack of a normal distribution here a problem??
fs.2pl<-factor.scores(fit.2pl)
#plot(fs.2pl)

#ICC and informaction curves
par(lwd=2)
plot(fit.2pl, legend = TRUE, cx = "left", cex.main = 1.5, cex.lab = 1.5, cex = 1.5)
plot(fit.2pl, type = "IIC", annot = FALSE, cex.main = 1.5, cex.lab = 1.5)
plot(fit.2pl, type = "IIC", items = 0, cex.main = 1.5, cex.lab = 1.5)

item.fit(fit.2pl, G = 4, FUN = mean)

```


####Analysis with 'mirt' package

Since we need to include the survey weights in this analysis, it was decided to use the mirt package, which allows survey weights. 

```{r}

#Rasch model
fit.r <- mirt(items2, model=1, itemtype='Rasch')
fit.rw <- mirt(items2, model=1, itemtype='Rasch', survey.weights = dat$weighth)

#2Pl model
fit.2 <- mirt(items2, model=1, itemtype='2PL')
fit.2w <- mirt(items2, model=1, itemtype='2PL', survey.weights = dat$weighth)

#Extract the discrimination parameters
ext.coef<-function(model, k) {
a <- coef(model, as.data.frame=T)
a[seq(k,length(a),4)]
a[c(1:8)]
}

qplot(x=ext.coef(fit.r,2), y=ext.coef(fit.rw,2), main="Difficulty parameters Rasch", xlab="No weights", ylab="Weights") + geom_line()
qplot(x=ext.coef(fit.2,1), y=ext.coef(fit.2w,1), main="Discrimination parameters 2PL", xlab="No weights", ylab="Weights") + geom_line()


```


```{r}

#Comparing discrimination parameter using ltm vs mirt (no weighting)
#First we fit the 2PL model using the mirt package
fit.2pl.mirt <- mirt(items2, model=1, itemtype='2PL')
coef(fit.2pl.mirt, simplify=T)

#Using mirt and survey weights
fit.2pl.mirtw <- mirt(items2, 1, itemtype='2PL', survey.weights = dat$weighth)

fit.2pl
fit.2pl.mirt
fit.2pl.mirtw

disc.ltm <- c(6.3415, 6.3599, 4.5101, 7.3436, 8.3184, 6.7368, 5.4509, 5.0711)
disc.mirt <- c(6.887, 6.878, 4.729, 7.78, 8.715, 7.052, 5.56, 5.163)
disc.mirtw <- c(7.031, 7.322, 4.856, 7.790, 8.616, 7.009, 5.536, 5.134)

qplot(x=disc.ltm, y=disc.mirt, main="Discrimination par압", xlab="ltm", ylab="mirt") + geom_line()

qplot(x=disc.mirt, y=disc.mirtw, main="Discrimination par압", xlab="No survey weights", ylab="Using survey weights") + geom_line()

```

#### Aalysis with weights

```{r}

fit.rasch.final <- mirt(items2, model=1, itemtype='Rasch', survey.weights = dat$weighth)
fit.2pl.final <- mirt(items2, model=1, itemtype='2PL', survey.weights = dat$weighth)

anova(fit.rasch.final, fit.2pl.final)  

#ICC압
plot(fit.rasch.final, type = 'trace')
plot(fit.2pl.final, type = 'trace')

```

